Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-16 p2/def-commands.tex
Improvements

Notice, however, that the commands listed above also encode some information about 
their input, the state of the filesystem
before the command is applied. In particular, \textit{create}$(n,\valv)$ requires that there are no files
or directories at $n$, while \textit{edit}$(n,\valv)$ and \textit{remove}$(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information available about their output than about their input.
As, based on the above, we expect that encoding more information in the commands
results in a model with greater predictive powers,
and in order to resolve this asymmetry, 
we propose a set of commands that encode
the type of the input value $\FS(n)$ as well.
(Some real-life filesystem commands like \textit{rmdir} do this already.)
Because the success of failure of commands depends on types of values,
it is not necessary to encode the actual input value.

@@

We model filesystem commands as partial functions 
that map filesystems to filesystems, but which
are only defined where the commands succeed.
As usual, if the function is not defined, we say that that it returns the bottom element, $\fsbroken$,
or that it \emph{breaks} the filesystem.
We describe a filesystem commands using its input, output, and
the node it is applied to.

\begin{mydef}[Filesystem commands]
Filesystem commands are partial endofunctions on filesystems,
which are not defined on filesystems where they return an error.
The commands are represented using triplets of the form
\[ \alpha = \caaaa{x}{y}{n}, \]
where $x,y\in\setv$ and $n\in\setn$.
The equivalence class $[x]$ is the \emph{input type} of the function;
$y$ is its \emph{output value} implicitly specifying the \emph{output type} $[y]$,
and $n$ is the node the command is applied to.
\end{mydef}
For example, $\cbfa{n}$ represents \textit{create}$(n,\valf)$,
and $\cdba{n}$ represents \textit{rmdir}$(n)$.

@@

\begin{mydef}[Effect of commands]
The effect of the command $\cxynv$ is defined as follows:
\[
\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\textrm{if $[\FS(n)]\neq[x]$,}\\
   \fsbroken &\textrm{if $\fsreplacement{\FS}{n}{y}$ violates the tree property,}\\
   \fsreplacement{\FS}{n}{y} &\textrm{otherwise.}
   \end{cases}
\]
\end{mydef}
In other words, the command $\cxynv$ breaks a filesystem if the input type $[x]$
does not match the type of the value in the filesystem at $n$, or if after replacing
the value at $n$ with $y$, the resulting filesystem ceases to satisfy the tree property.

\myskip
Depending on their input and output types, there are nine types of commands,
which we separate into four categories depending on the overall effect of the command.
Constructions commands extend the filesystem,
while destruction commands shrink it;
the replacement command replaces a file value with a file value
(which may or may not be different from the original value),
and assertion commands either break the filesystem, or assert the type of the value at a node.
\begin{mydef}[Command categories]\mlabel{def:command_categories}
Depending on their input and output types, filesystem commands
belong to exactly one of the following four categories:
\begin{itemize}
\item[]\emph{Construction commands:}
    $\cbfa{n}$, $\cbda{n}$ and $\cfda{n}$;
\item[]\emph{Destruction commands:}
    $\cdfa{n}$, $\cdba{n}$ and $\cfba{n}$;
\item[]\emph{Assertion commands:}
    $\cbba{n}$ and $\cdda{n}$;
\item[]\emph{Replacement command:}
    $\caaa{\valfx}{\valff}{n}$.
\qedhere
\end{itemize}
\end{mydef}

@@

\myskip
For reasons also listed in \cite{NREC}, in this model we will not consider
a \textit{move} or \textit{rename} command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for \textit{move} would call for a more complicated model.
As \cref{sec_multidir} details, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by post-processing changes in the filesystem to
discover them, which can enhance usability.

>>>>

Notice, however, that the commands listed above also encode some information about 
their input, the state of the filesystem
before the command is applied. In particular, \textit{create}$(n,\valv)$ requires that there are no files
or directories at $n$, while \textit{edit}$(n,\valv)$ and \textit{remove}$(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information available about their output than about their input.
As, based on the above, we expect that encoding more information in the commands
results in a model with greater predictive powers,
and in order to resolve this asymmetry, 
we propose a set of commands that encode
the type of the input value $\FS(n)$ as well.
(Some real-life filesystem commands like \textit{rmdir} do this already.)
Because the success or failure of commands depends on types of values,
it is not necessary to encode the actual input value.

@@

According to the above,
we model filesystem commands with partial functions 
that map filesystems to filesystems, but which
are only defined where the commands succeed.
As usual, if the function is not defined, we say that that it returns the bottom element, $\fsbroken$,
or that it \emph{breaks} the filesystem.
We describe a filesystem commands using its input, output, and
the node it is applied to.

\begin{mydef}[Filesystem commands]
Filesystem commands are partial endofunctions on filesystems,
which are not defined on filesystems where they return an error.
The commands are represented using triplets of the form
\[ \alpha = \caaaa{x}{y}{n}, \]
where $x,y\in\setv$ and $n\in\setn$.
The equivalence class $[x]$ is the \emph{input type} of the function;
$y$ is its \emph{output value} implicitly specifying the \emph{output type} $[y]$,
and $n$ is the node the command is applied to.
\end{mydef}
For example, $\cbfa{n}$ represents \textit{create}$(n,\valfx)$,
and $\cdba{n}$ represents \textit{rmdir}$(n)$.

@@

\begin{mydef}[Effect of commands]
The effect of the command $\cxynv$ is defined as follows:
\[
\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\textrm{if $[\FS(n)]\neq[x]$,}\\
   \fsbroken &\textrm{if $\fsreplacement{\FS}{n}{y}$ violates the tree property,}\\
   \fsreplacement{\FS}{n}{y} &\textrm{otherwise.}
   \end{cases}
\]
\end{mydef}
In other words, the command $\cxynv$ breaks a filesystem if the input type $[x]$
does not match the type of the value in the filesystem at $n$, or if after replacing
the value at $n$ with $y$, the resulting filesystem ceases to satisfy the tree property
in \cref{def_treeprop}.

\myskip
There are nine groups of commands considering their input and output types.
Based on these groups we separate commands into four categories 
that reflect their overall effect.
Constructions commands extend the filesystem,
while destruction commands shrink it;
the replacement command replaces a file value with a file value
(which may or may not be different from the original value),
and assertion commands either break the filesystem, or assert the type of the value at a node.
\begin{mydef}[Command categories]\mlabel{def:command_categories}
Depending on their input and output types, filesystem commands
belong to exactly one of the following four categories:
\begin{itemize}
\item[]\emph{Construction commands:}
    $\cbfa{n}$, $\cbda{n}$ and $\cfda{n}$;
\item[]\emph{Destruction commands:}
    $\cdfa{n}$, $\cdba{n}$ and $\cfba{n}$;
\item[]\emph{Assertion commands:}
    $\cbba{n}$ and $\cdda{n}$;
\item[]\emph{Replacement command:}
    $\caaa{\valfx}{\valff}{n}$.
\qedhere
\end{itemize}
\end{mydef}

@@

\myskip
For reasons also listed in \cite{NREC}, in this model we will not consider
a \textit{move (rename)} command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for \textit{move} would call for a more complicated model.
As \cref{sec_multidir} details, even with this restriction,
an implementation of the reconciliation algorithm presented here
can still handle \emph{move} commands in its input and output
by splitting them into \emph{delete} and \emph{create} commands,
and merging them later.

### 2018-07-16 p2/def-filesystem.tex
Improvements

\begin{mydef}[Filesystem values: $\vald$, $\valfx$, $\empt$, $\eqclass{v}$]
The set of values $\setv$ is partitioned into directories, files, and the empty value $\empt$.
As usual, we write $[v]$ for the the equivalence class of $v\in\setv$ according to this partition,
which represents the type of the value.
File and directory values are denoted by $\valfx$ and $\vald$, respectively,
with or without indices, and so the partitioning is in fact $[\vald], [\valfx], [\empt]$.
\end{mydef}

@@

Every filesystem must have a so-called \emph{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.
Using the notation introduced above, we can formally express this as follows.
\begin{mydef}[Tree property]
A filesystem $\FS$ has the tree property iff
\[ \forall n\in\setn:
\FS(n) \neq \empt \: \Longrightarrow \: \Big[\FS\big(\parentf{n}\big)\Big] = [\vald] \]
wherever $\parent(n)$ is defined, that is, $n$ has a parent.
\end{mydef}

@@

An essential additional assumption in our model is that there is only
a \emph{single directory value} ($|\setd|=1$).
The realization that
this creates a symmetry between directories and empty nodes
and makes it possible to exploit a hidden symmetry in filesystem commands
that allows formulating a correct and complete reconciliation algorithm
is one of the main contributions of this paper.
In \cref{sec_multidir} we explore the applicability of our model given this assumption:
why applicability may not be affected and ways to relax this restriction.

>>>>

\begin{mydef}[Filesystem values: $\vald$, $\valfx$, $\empt$, $\eqclass{v}$]
The set of values $\setv$ is partitioned into directories, files, and the empty value $\empt$.
As usual, we write $[v]$ for the the equivalence class of $v\in\setv$ according to this partition,
which represents the type of the value.
Specific directory and file values are denoted by $\vald$ and $\valfx$, respectively,
and so the partitioning is in fact $[\vald], [\valfx], [\empt]$.
\end{mydef}

@@

Every filesystem must have a so-called \emph{tree property}, which means that
if the filesystem is not empty at a node, and the node has a parent,
then there must be a directory at the parent node.
Using the notation introduced above, we can formally express this as follows.
\begin{mydef}[Tree property]\mlabel{def_treeprop}
A filesystem $\FS$ has the tree property iff
\[ \forall n\in\setn:
\FS(n) \neq \empt \: \Longrightarrow \: \Big[\FS\big(\parentf{n}\big)\Big] = [\vald] \]
wherever $\parent(n)$ is defined, that is, $n$ has a parent.
\end{mydef}

@@

An essential additional assumption in our model is that there is only
a \emph{single directory value} ($|\setd|=1$).
The realization that
this creates a symmetry between directories and empty nodes
and makes it possible to exploit a hidden symmetry in filesystem commands
that allows formulating a correct and complete reconciliation algorithm
is one of the main contributions of this paper.
In \cref{sec_multidir} we explore the applicability of our model given this assumption;
specifically,
why applicability may not be affected and ways to relax this restriction.

We also assume that there are multiple different file contents, that is, $|[\valfx]|>1$.
